Measuring Political Freedom Around the World

### 2020-11-13 slides.Rmd
added textbook quote
added plots getting there

```{r setup, include=FALSE}
source(here::here("R", "regression.R"))
theme_set(theme_light(base_size = 30))
options(htmltools.dir.version = FALSE, echo = FALSE)
knitr::opts_chunk$set(echo = FALSE, fig.width=12, fig.height=8) 
```

@@

@@

@@

```{r message=FALSE}
data %>% 
  group_by(country) %>% 
summarise(across(.cols = c(corruption_perception_index_cpi:murder_per_mil_people), mean)) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  kable(
    col.names = c("Country", "CPI", "Edu Exp", "Electric Use", "Gini", "Internet Users", "Labor Force", "Life Expectancy","Military Spending", "Murder Rate")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 6, fixed_thead = T)
```

@@

  geom_text(aes(x = polrights_fh, y = n, label = n), vjust = -.75) + 
  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
  scale_x_continuous(breaks = 1:7) + 
  theme_test(base_size = 30) + 
  theme(legend.position="none")
```

@@

# Best Subsets by Adjusted R Squared
```{r message=FALSE}
subsets %>% 
  group_by(df) %>% 
  filter(adj.r.squared == max(adj.r.squared)) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  rename(a_rsq = adj.r.squared) %>% 
  select(df, a_rsq, model) %>% 
  arrange(desc(a_rsq)) %>% 
DT::datatable(fillContainer = FALSE, options = list(pageLength = 3), 
              colnames = c("DF", "A_Rsq", "Model"))
```

@@

```{r include=FALSE}
equatiomatic::extract_eq(model = model_df5, intercept = "beta", terms_per_line = 1)
```

@@

```{r}
broom::tidy(model_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names = c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) 
```

***
```{r}
broom::glance(model_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(model_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"))  %>% 
  row_spec(2, color = "white", background = "#D7261E")
```

@@

```{r}
broom::glance(model_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(model_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(3, color = "white", background = "#D7261E")
```

@@

```{r}
broom::glance(model_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(model_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(4, color = "white", background = "#D7261E")
```

@@

```{r}
broom::glance(model_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

@@

class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(model_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(5, color = "white", background = "#D7261E")
```

@@

```{r}
broom::glance(model_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(model_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(6, color = "white", background = "#D7261E")
```

@@

@@

@@

```{r}
model_df5 %>% 
  broom::augment() %>% 
  mutate(yhat = predict(model_df5),
         resid = resid(model_df5)) %>% 
  ggplot() + 
  geom_point(aes(x = yhat, y = resid)) + 
  geom_hline(yintercept = 0, color = "red") + 
  scale_y_continuous(limits = c(-4,4))
```

@@

* plot
* 
* vif
car::vif(fit)

@@

@@

<font size="2">
```{r results='asis'}
stargazer::stargazer(fit_df6, model_df5, fit_df4, type = "html", 
                     title = "Regression Results", 
                     covariate.labels = c(),
                     header = FALSE,
                     single.row = TRUE)
```

>>>>

```{r setup, include=FALSE}
source(here::here("R", "regression.R"))
theme_set(theme_light(base_size = 30))
options(htmltools.dir.version = FALSE, echo = FALSE)
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 8) 
```

@@

---
# OECD

@@

Inputing median

@@

---
# needs work
```{r message=FALSE}
data %>% 
  group_by(country) %>% 
summarise(across(.cols = c(corruption_perception_index_cpi:murder_per_mil_people), mean)) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  kable(
    col.names = c("Country", "CPI", "Edu Exp", "Electric Use", "Gini", "Internet Users", "Labor Force", "Life Expectancy","Military Spending", "Murder Rate")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 6, fixed_thead = T)
```

@@

  geom_text(aes(x = polrights_fh, y = n, label = n), vjust = -.75) + 
  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
  scale_x_continuous(breaks = 1:7) + 
  theme_test(base_size = 30) + 
  theme(legend.position = "none")
```

@@

# Best Subsets by Adjusted R Squared
?latex
```{r message=FALSE}
subsets %>% 
  group_by(df) %>% 
  filter(adj.r.squared == max(adj.r.squared)) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  rename(a_rsq = adj.r.squared) %>% 
  select(df, a_rsq, model) %>% 
  arrange(desc(a_rsq)) %>% 
DT::datatable(fillContainer = FALSE, options = list(pageLength = 3), 
              colnames = c("DF", "A_Rsq", "Model"))
```

@@

```{r include=FALSE}
equatiomatic::extract_eq(model = fit_df5, intercept = "beta", terms_per_line = 1)
```

@@

```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names = c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) 
```

***
```{r}
broom::glance(fit_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"))  %>% 
  row_spec(2, color = "white", background = "#D7261E")
```

@@

```{r}
broom::glance(fit_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(3, color = "white", background = "#D7261E")
```

@@

```{r}
broom::glance(fit_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

---
class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(4, color = "white", background = "#D7261E")
```

@@

```{r}
broom::glance(fit_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

@@

class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(5, color = "white", background = "#D7261E")
```

@@

```{r}
broom::glance(fit_df5) %>% 
  select(r.squared, adj.r.squared, adj.r.squared, AIC, BIC) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable()
```

class: center
$$
\small
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$
```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  row_spec(6, color = "white", background = "#D7261E")
```

@@

plots resids on cov
---

@@

```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"),
        caption = "Included Outliers") 
```

```{r}
data %>% 
  filter(polrights_fh >= 4) %>% 
  lm(model_df5_formula, .) %>% 
  broom::tidy() %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"),
        caption = "Removed Outliers") 
```

```{r}
broom::tidy(fit_df5) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"),
        caption = "Included Outliers") %>% 
  row_spec(3, color = "white", background = "#D7261E")
```

```{r}
data %>% 
  filter(polrights_fh >= 4) %>% 
  lm(model_df5_formula, .) %>% 
  broom::tidy() %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value)) %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value"),
        caption = "Removed Outliers") %>% 
  row_spec(3, color = "white", background = "#D7261E")
```

# Discussion of outliers

- We identified these outliers.
- confirmed they are not errors. 
- decided to keep them because They are indicating a substancial loss in political freedom which falls outside of the scope of the model. However, is important for our reserach Q.
***
> A safe rule frequently suggested is to discard an outlier only if there is direct evidence that it represents an error in recording, a miscalculation, a malfunctioning of equipment, or a similar type of circumstance. -Chapter 3 p.108

@@

```{r}
fit_df5 %>% 
  broom::augment() %>% 
  mutate(yhat = predict(fit_df5),
         resid = resid(fit_df5)) %>% 
  ggplot() + 
  geom_point(aes(x = yhat, y = resid)) + 
  geom_hline(yintercept = 0, color = "red") + 
  scale_y_continuous(limits = c(-4,4))
```

```{r}
car::vif(fit_df5)
```
* vif

@@

@@

1. we think this
1. because

@@

<font size="2">
```{r results='asis'}
stargazer::stargazer(fit_df6, fit_df5, fit_df4, fit_df3, type = "html", 
                     title = "Regression Results", 
                     covariate.labels = c(),
                     header = FALSE,
                     single.row = TRUE)
```
